---
title: "Debugging boot()"
author: "Esa Turkulainen"
date: "4/22/2022"
output: html´_document
---

```{r setup, include = FALSE}
knitr::opts_knit$set(root.dir = "~/CRP_enrichment") # set working directory
library(tidyverse)
library(boot)
```

```{r load_data, echo = FALSE}
# .rdata by Sofie Ekroos, 2021

# Load data on individual donations
load("~/CRP_enrichment/data/r02.fd.bd.all.rdata") # outputs an object called "output" into the environment
donations <- output

# Load FinDonor demographic data
load("~/CRP_enrichment/data/r02ds.donorData.rdata") # outputs an object called "output" into the environment
findonor <- output 

# Load THL data
# Sofie: thldalta.rdata contains all five THL cohorts, extract FINRISK97 and Health2000 from the others
load("~/CRP_enrichment/data/thldata.rdata")
fr1997 <- thldata$fr1997
h2000 <- thldata$h2000

# Remove leftovers
rm(output)
rm(thldata)

# We only want to look at first donation event values from each donor
donors <- donations %>%
    group_by(donor) %>%
    filter(date == min(date)) %>%
    ungroup()
```

# Boot funcs

```{r}
bootmean <- function(data, i) {
    splice <- data[i]
    return(mean(splice, na.rm = T))
}

bootmedian <- function(data, i) {
    splice <- data[i]
    return(median(splice, na.rm = T))
}

get_ratio <- function(cohort, var1, var2, var1_trld, var2_trld) {
    # Answers the question:
    # How much the proportion of individuals with var2 >< var2_trld
    # increase after we filter the population with var1_trld

    # Base population: cohort
    vi <- na.omit(cohort[, c(var1, var2)])

    # Find individuals of interest (ioi, above/below of var2_trld)
    ioi <- vi[vi[, var2] >= var2_trld, ]

    # Get proportion
    prop <- 100 * nrow(ioi) / nrow(vi)

    # Now filter base population first using var1
    filtered <- vi[vi[, var1] >= var1_trld, ]

    # And find IoI again
    f_ioi <- filtered[filtered[, var2] >= var2_trld, ]

    # Now get proportion (filtered population)
    f_prop <- 100 * nrow(f_ioi) / nrow(filtered)

    # Compute difference between proportions
    diff <- round(f_prop - prop, 2)

    # Return
    return(diff)
}

get_ratio_boot <- function(cohort, var1, var2, var1_trld, var2_trld, i) {
    # Answers the question:
    # How much the proportion of individuals with var2 >< var2_trld
    # increase after we filter the population with var1_trld

    # Base population: cohort
    vi <- na.omit(cohort[i, c(var1, var2)])
    
    # Find individuals of interest (ioi, above/below of var2_trld)
    ioi <- vi[vi[, var2] >= var2_trld, ]

    # Get proportion
    prop <- 100 * nrow(ioi) / nrow(vi)

    # Now filter base population first using var1
    filtered <- vi[vi[, var1] >= var1_trld, ]

    # And find IoI again
    f_ioi <- filtered[filtered[, var2] >= var2_trld, ]

    # Now get proportion (filtered population)
    f_prop <- 100 * nrow(f_ioi) / nrow(filtered)

    # Compute difference between proportions
    diff <- round(f_prop - prop, 2)

    # Return
    return(diff)
}
```

# Testing

Bootstrapping should always converge to the "population" estimate. When we bootstrap, we consider the sample space to be our "population". So, if check the mean of a varible first by taking the sample mean

```{r sample_mean}
mean(h2000$GP, na.rm = T)
```
we should get roughly the same number by using our bootmean() statistic with boot().

```{r bootmean}
boot(h2000$GP, statistic = bootmean, R = 100)$t0
```
Let's test this with the median now:

```{r sample_median}
median(h2000$GP, na.rm = T)
```

```{r bootmedian}
boot(h2000$GP, statistic = bootmedian, R = 100)$t0
```
These seem to work fine. Next, we want to check the proportion of certain kinds of people in a population before and after applying some kind of variable filter. For example, the proportion of people above median GP before and after we filter for ferritin lower bound of 5 ug/l (so, the difference).

This is the result we got straight from our sample:

```{r sample_prop}
get_ratio(h2000, var1 = "FERRITIINI", var2 = "GP", var1_trld = 5, var2_trld = 1.16)
```
And this is what comes out of boot:

```{r boot_prop}
boot(h2000, statistic = get_ratio_boot, R = 1000, var1 = "FERRITIINI", var2 = "GP", var1_trld = 5, var2_trld = 1.16)$t0
```

With this configuration of the get_ratio function the boot() seems to work correctly. Can we break it in a surprising way? (We know we can, this is just a demo).


```{r get_ratio_broken_index}

get_ratio_broken_index <- function(cohort, var1, var2, var1_trld, var2_trld, i) {
    
    # Base population: cohort
    vi <- cohort %>%
        select({{var1}}, {{var2}}) %>%
        drop_na()

    splice <- vi[i, ]
    
    # Find individuals of interest (ioi, above/below of var2_trld)
    ioi <- splice %>%
        filter({{var2}} >= {{var2_trld}})

    # Get proportion
    prop <- 100 * nrow(ioi) / nrow(splice)

    # Now filter base population first using var1
    filtered <- splice %>%
        filter({{var1}} >= {{var1_trld}})

    # And find IoI again
    f_ioi <- filtered %>% 
        filter({{var2}} >= {{var2_trld}})

    # Now get proportion (filtered population)
    f_prop <- 100 * nrow(f_ioi) / nrow(filtered)

    # Compute difference between proportions
    diff <- round(f_prop - prop, 2)

    # Return
    return(diff)
}

```

```{r broken_index_test}
boot(h2000, statistic = get_ratio_broken_index, R = 1, var1 = FERRITIINI, var2 = GP, var1_trld = 5, var2_trld = 1.16)$t0
```
Moving the index splicing after the selection and NA removal breaks the result. I suspect this is because the length(i) = nrow(h2000), and length(i) != nrow(h2000 %>% select(var1, var2) %>% drop_na()). Let's see:

```{r get_ratio_broken_index_nodrop}

get_ratio_broken_index_nodrop <- function(cohort, var1, var2, var1_trld, var2_trld, i) {

    # Base population: cohort
    vi <- cohort %>%
        select({{var1}}, {{var2}})
    
    splice <- vi[i, ]
    
    # Find individuals of interest (ioi, above/below of var2_trld)
    ioi <- splice %>%
        filter({{var2}} >= {{var2_trld}})

    # Get proportion
    prop <- 100 * nrow(ioi) / nrow(splice)

    # Now filter base population first using var1
    filtered <- splice %>%
        filter({{var1}} >= {{var1_trld}})

    # And find IoI again
    f_ioi <- filtered %>% 
        filter({{var2}} >= {{var2_trld}})

    # Now get proportion (filtered population)
    f_prop <- 100 * nrow(f_ioi) / nrow(filtered)

    # Compute difference between proportions
    diff <- round(f_prop - prop, 2)

    # Return
    return(diff)
}

```

```{r nodrop_test}
boot(h2000, statistic = get_ratio_broken_index_nodrop, R = 1, var1 = FERRITIINI, var2 = GP, var1_trld = 5, var2_trld = 1.16)$t0
```

Curious! It's still wrong. Does select() change the size of the data so that the indexing fails somehow?

```{r}
nrow(h2000)

selected <- h2000 %>% select(FERRITIINI, GP)
nrow(selected)
```
No, it doesn't (as it shouldn't). Let's go back to the original working one to see what happens with the indexing.


```{r}
get_ratio_boot_i <- function(cohort, var1, var2, var1_trld, var2_trld, i) {
    # Answers the question:
    # How much the proportion of individuals with var2 >< var2_trld
    # increase after we filter the population with var1_trld

    # Base population: cohort
    vi <- na.omit(cohort[, c(var1, var2)])
    
    vi <- vi[i, ]
    
    # Find individuals of interest (ioi, above/below of var2_trld)
    ioi <- vi[vi[, var2] >= var2_trld, ]

    # Get proportion
    prop <- 100 * nrow(ioi) / nrow(vi)

    # Now filter base population first using var1
    filtered <- vi[vi[, var1] >= var1_trld, ]

    # And find IoI again
    f_ioi <- filtered[filtered[, var2] >= var2_trld, ]

    # Now get proportion (filtered population)
    f_prop <- 100 * nrow(f_ioi) / nrow(filtered)

    # Compute difference between proportions
    diff <- round(f_prop - prop, 2)

    # Return
    return(diff)
}
```

```{r broken_test}
boot(h2000, statistic = get_ratio_boot_i, R = 100, var1 = "FERRITIINI", var2 = "GP", var1_trld = 5, var2_trld = 1.16)$t0
```
We get really close, but still incorrect. Even though our approach is identical to the tidyverse version, we get a different result (tidy: 8, this: 0.2, correct: 0.1). We'll need to make this simpler to understand what happens.

```{r}
# Let's create a dataset with controlled NA (1000 pcs)

vi <- na.omit(h2000[, c("FERRITIINI", "GP")])
scuffed_data <- vi
set.seed(2022)
scuffed_data[sample(1:nrow(vi), 1000, replace = F), ] <- NA

# Now let's see if we get identical results from our two different implementations
var1 <- "FERRITIINI"
var2 <- "GP"
var1_trld <- 5
var2_trld <- 1.16
## BASE

# Find individuals of interest (ioi, above/below of var2_trld)
ioi_base <- scuffed_data[scuffed_data[, var2] >= var2_trld, ]
# Get proportion
prop_base <- 100 * nrow(ioi_base) / nrow(scuffed_data)
# Now filter base population first using var1
filtered_base <- scuffed_data[scuffed_data[, var1] >= var1_trld, ]
# And find IoI again
f_ioi_base <- filtered_base[filtered_base[, var2] >= var2_trld, ]
# Now get proportion (filtered population)
f_prop_base <- 100 * nrow(f_ioi_base) / nrow(filtered_base)
# Compute difference between proportions
diff_base <- round(f_prop_base - prop_base, 2)

## TIDY
# Find individuals of interest (ioi, above/below of var2_trld)
ioi_tidy <- scuffed_data %>%
    filter(GP >= var2_trld)

# Get proportion
prop_tidy <- 100 * nrow(ioi_tidy) / nrow(scuffed_data)

# Now filter base population first using var1
filtered_tidy <- scuffed_data %>%
    filter(FERRITIINI >= var1_trld)

# And find IoI again
f_ioi_tidy <- filtered_tidy %>% 
    filter(GP >= var2_trld)

# Now get proportion (filtered population)
f_prop_tidy <- 100 * nrow(f_ioi_tidy) / nrow(filtered_tidy)

# Compute difference between proportions
diff_tidy <- round(f_prop_tidy - prop_tidy, 2)


print(c(diff_base, diff_tidy))
```
We can now see that they produce wildly different estimates. Tidy must handle NAs differently here. Let's see how.

```{r}
nrow(ioi_base) 
nrow(ioi_tidy)
```
We find that filter() strips NA automatically (which you can also find by reading the documentation). This in turn affects the proportion calculations, as it is done using nrow(). I consider this solved.
