---
title: "CRP_OC_investigation"
author: "Esa"
output: 
    html_document:
        keep_md: yes
---

```{r}
#| label: setup
#| include: false
knitr::opts_knit$set(root.dir = "~/CRP_enrichment") # set working directory
source("~/CRP_enrichment/src/funcs.R") # load our helper functions
library(tidyverse)
### For prettier plots
library(ggridges)
library(viridis)
library(ggpubr)
### for bootstrapping
library(boot)
### for easy exclusion table: https://github.com/entjos/ExclusionTable
library(ExclusionTable)

boot_n <- 10000 # number of bootstrap samples
ferritin_values <- seq(0, 50, 1) # Start filtering thresholds from 0 (no filtering / current situation)
iterations <- length(ferritin_values)
CRP_trld <- 3
set.seed(01062022) # set seed
```


In this additional document we investigate what happens when we replicate the analysis, but considering the use of oral contraceptives. 

# Create datasets

```{r}
#| label: create_datasets
#| code-summary: Create datasets
#| message: false
#| warning: false

# Load data on individual donations
load("~/CRP_enrichment/data/r02.fd.bd.all.rdata") # outputs an object called "output" into the environment
donations <- output

# We only want to look at first donation event values from each donor
donors <- donations %>%
    group_by(donor) %>%
    filter(date == min(date)) %>%
    ungroup()

# Load FinDonor demographic data
load("~/CRP_enrichment/data/r02ds.donorData.rdata") # outputs an object called "output" into the environment
findonor <- output

# Combine the FinDonor datasets
FinDonor <- left_join(donors, findonor, by = "donor")

# Load THL data
# Sofie: thldalta.rdata contains all five THL cohorts, extract FINRISK97 and Health2000 from the others
load("~/CRP_enrichment/data/thldata.rdata")
FR97 <- thldata$fr1997
H2000 <- thldata$h2000

# Remove leftovers
rm(output)
rm(thldata)

## Rename useful stuff
# Ferritin, Self-Reported Health
FinDonor <- rename(FinDonor, SRH = QR17, Menstruation = QR79, Age_float = Age, Age = age)
FR97 <- rename(FR97, Ferritin = FERRITIN, SRH = Q40, Gender = SUKUP, Menstruation = K129, Age = IKA)
H2000 <- rename(H2000, Ferritin = FERRITIINI, SRH = BA01, Gender = SP2, Menstruation = BD03, Age = IKA2, Menopause = MENOP, APOB = NMR_APOB, APOA1 = NMR_APOA1)

# Make "useful stuff" conform with each other
FinDonor <- FinDonor %>% 
    mutate(SRH = case_when(SRH == "Excellent" ~ 1,
                           SRH == "Very_good" ~ 2,
                           SRH == "Good" ~ 3,
                           SRH == "Satisfactory" ~ 4,
                           SRH == "Poor" ~ 5),
           Group = case_when(Gender == "Men" ~ "Men",
                             Gender == "Women" & (Menstruation == "regular_period" | Menstruation == "irregular_period") ~ "Women|Menstr",
                             Gender == "Women" & Menstruation == "no_period" ~ "Women|Non-menstr",
                             TRUE ~ "NA")) # Equates to "else"

FR97 <- FR97 %>%
    mutate(Gender = case_when(Gender == 1 ~ "Men",
                              Gender == 2 ~ "Women",
                              TRUE ~ "NA"),
           Group = case_when(Gender == "Men" ~ "Men",
                             Gender == "Women" & (Menstruation == 1 | Menstruation == 2) ~ "Women|Menstr",
                             Gender == "Women" & Menstruation == 3 ~ "Women|Non-menstr",
                             TRUE ~ "NA"),
           # THIS GROUP IS FOR ORAL CONTRACEPTION USE, ADDED DURING REVIEW
           Hormonal = case_when(Gender == "Men" ~ "Men",
                                Gender == "Women" & KY151 == 1 ~ "Yes",
                                Gender == "Women" & KY151 != 1 ~ "No",
                                TRUE ~ "NA"))

H2000 <- H2000 %>%
    mutate(Gender = case_when(Gender == 1 ~ "Men",
                              Gender == 2 ~ "Women",
                              TRUE ~ "NA"),
           # It is worth noting here, that menstruation status was not asked from women over 55. So, we treat these as postmenopausal.
           Group = case_when(Gender == "Men" ~ "Men",
                             Gender == "Women" & (Menstruation == 1 | Menstruation == 2) ~ "Women|Menstr",
                             Gender == "Women" & (Menstruation == 3 | Age >= 55) ~ "Women|Non-menstr",
                             TRUE ~ "NA"),
            # THIS GROUP IS FOR ORAL CONTRACEPTION USE, ADDED DURING REVIEW
           Hormonal = case_when(Gender == "Men" ~ "Men",
                                Gender == "Women" & BD26 == 1 ~ "Yes",
                                Gender == "Women" & BD26 == 0 ~ "No",
                                TRUE ~ "NA"))

# Load in additional H2000 variable BA14: Stroke
newdata <- read_tsv("~/CRP_enrichment/data/THLBB2020_19_T2000_lisapoiminta_10102022.txt")
strokedata <- newdata %>% 
    select(RELEASE_ID, BA14)
# Merge with H2000
H2000 <- merge(H2000, strokedata, by = "RELEASE_ID", all.x = TRUE)

# Donation eligibility
# These are both "approximates" in a sense, we don't have all the necessary variables to
# filter thoroughly, and we'll be able to do more filtering on Health2000 than FR97
donor_eligible_h2k <- H2000 %>% 
    filter(BMII_PAINO.x >= 50 & BMII_PAINO.x <= 200) %>% # Filter away people <50kg and >200kg
    filter(Age >= 18 & Age <= 66) %>% # Filter away too young and too old
    filter((B_Hb >= 125 & Gender == "Women") | (B_Hb >= 135 & Gender == "Men")) %>% # Filter by hemoglobin
    filter(BA08 == 0) %>% # filter out people with heart attacks
    filter(BA09 == 0) %>% # filter out people with angina
    filter(BA10 == 0) %>% # cardiac insufficiency / heart failure
    filter(BA14 == 0) %>% # Stroke
    filter(!(BA26 == 1 & ATC_A10A == 1)) %>% # filter out people who are diabetic AND use insulin
    filter(SRH < 4) %>% # filter out "Bad" or "Very bad" SRH
    rename(GlycA = GP) %>% # rename for ease of use
    mutate(HbA1C = B_GHb_A1C * 10.93 - 23.50)

donor_eligible_fr <- FR97 %>%
    filter(PAINO >= 50 & PAINO <= 200) %>% # Filter away people <50kg and >200kg
    filter(Age >= 18 & Age <= 66) %>% # Filter away too young and too old
    #filter((HGB >= 125 & Gender == 2) | (HGB >= 135 & Gender == 1)) %>% # DON'T filter by hemoglobin, < 500 values in data
    filter(Q15A != 2) %>% # STEMI, NSTEMI
    filter(Q16A != 2) %>% # Stroke
    filter(is.na(Q38) | Q38 == 1 | Q38 == 3) %>%  # Insulin treatment (2: just insulin, 4: insulin and a tablet)
    filter(Q17B != 2) %>% # cardiac insufficiency
    filter(Q17C != 2) %>% # angina pectoris
    filter(SRH < 4) %>% # filter out "Bad" or "Very bad" SRH
    rename(GlycA = GP) # rename for ease of use

de_fr_omit_OC <- donor_eligible_fr %>% filter(Hormonal != "Yes")
de_h2k_omit_OC <- donor_eligible_h2k %>% filter(Hormonal != "Yes")

fer_crp <- bind_rows(FR97 = donor_eligible_fr[, c("Ferritin", "Group", "CRP")], 
                     H2000 = donor_eligible_h2k[, c("Ferritin", "Group", "CRP")], .id = "Cohort") %>% 
    mutate(Group = ordered(Group, levels = c("Women|Menstr", "Women|Non-menstr", "Men")),
           Cohort = ordered(Cohort, levels = c("FR97", "H2000"))) %>%
    filter(Group != "NA") %>%
    filter(CRP >= 0.01) %>%
    drop_na()
fer_crp_omit_OC <- bind_rows(FR97 = de_fr_omit_OC[, c("Ferritin", "Group", "CRP")], 
                     H2000 = de_h2k_omit_OC[, c("Ferritin", "Group", "CRP")], .id = "Cohort") %>% 
    mutate(Group = ordered(Group, levels = c("Women|Menstr", "Women|Non-menstr", "Men")),
           Cohort = ordered(Cohort, levels = c("FR97", "H2000"))) %>%
    filter(Group != "NA") %>%
    filter(CRP >= 0.01) %>%
    drop_na()
# fer_crp_nonhormonal <- bind_rows(FR97 = donor_eligible_fr[, c("Ferritin", "Hormonal", "CRP")], 
#                          H2000 = donor_eligible_h2k[, c("Ferritin", "Hormonal", "CRP")], .id = "Cohort") %>% 
#     mutate(Group = ordered(Hormonal, levels = c("No", "Yes", "Men")),
#            Cohort = ordered(Cohort, levels = c("FR97", "H2000"))) %>%
#     filter(Group != "NA") %>%
#     filter(CRP >= 0.01) %>%
#     drop_na()

# NB! Use of contraceptives is coded in
# FR97: KY151: "Käytättekö tällä hetkellä ehkäisypillereitä?" (1 yes 2 no but earlier yes 3 no never)
# FR97: KY152: "Käytättekö tällä hetkellä ehkäisyyn tavallista kierukkaa?" (1 yes 2 no but earlier yes 3 no never)
# FR97: KY153: "Käytättekö tällä hetkellä ehkäisyyn hormonikierukkaa?" (1 yes 2 no but earlier yes 3 no never)
# H2K: Lääkekysely ATC_G03A: Hormonaaliset ehkäisyvalmisteet (0 no 1 yes)
# H2K: BD26: "Käytättekö tällä hetkellä ehkäisypillereitä?" (0 no 1 yes)
# H2K: BD26a: "Kuinka monen vuoden ajan yhteensä olette käyttänyt ehkäisypillereitä?" (int)
# H2K: BD27: "Oletteko joskus aiemmin käyttänyt ehkäisypillereitä?" (0 never 1 yes)
# H2K: BD27a: Same as BD26a
# H2K: BD28a: "Käytättekö tällä hetkellä ehkäisyyn hormonikierukkaa?" (0 no 1 yes) 
# H2K: BD28b: "Käytättekö tällä hetkellä ehkäisyyn tavallista kierukkaa?" (0 no 1 yes)
# H2K: BD29: "Kuinka monen vuoden ajan yhteensä olette käyttänyt hormonikierukkaa?" (int)
# H2K: BD30a: "Oletteko joskus aiemmin käyttänyt ehkäisyyn hormonikierukkaa?" (0 never 1 yes)
# H2K: BD30b: "Oletteko joskus aiemmin käyttänyt ehkäisyyn tavallista kierukkaa?" (0 never 1 yes) 
# H2K: BD31: Same as BD29
```

```{r}
#| label: plot_CRP_by_hormonal_use
#| code-summary: CRP densities by cohort
#| warning: false
#| message: false

####
# HORMONAL CONTRACEPTION VS CRP
####

ggplot(data = donor_eligible_fr %>% filter(Gender == "Women" & Hormonal != "NA" & Age < 55), aes(x = CRP, group = Hormonal, fill = Hormonal)) +
    geom_density(adjust = 1.5, alpha = .4) +
    theme_minimal() +
    scale_x_log10() +
    labs(title = "FR97")

ggplot(data = donor_eligible_h2k %>% filter(Gender == "Women" & Hormonal != "NA" & Age < 55), aes(x = CRP, group = Hormonal, fill = Hormonal)) +
    geom_density(adjust = 1.5, alpha = .4) +
    theme_minimal() +
    scale_x_log10() +
    labs(title = "H2000")
```

We do see that the use of OC increases the CRP levels in the population of women under 55.

**But what happens with the actual analysis?**

## Original analysis
```{r ratio_boot_CRP_by_subgroup}
if (!file.exists(paste0("~/CRP_enrichment/data/PUBL_finrisk_CRP_", boot_n, ".rds"))) { # run bootstrap only if needed
    
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    bootobs <- setNames(vector("list", length = 3), c("men", "women_mens", "women_nonmens"))
    
    for (i in 1:iterations) {
        
        #############
        #### FR97
        #############
        
        ## Compute
        # Men
        bootobs$men[[i]] <- boot(fer_crp %>% filter(Group == "Men" & Cohort == "FR97"), statistic = get_ratio_boot, R = boot_n, 
                             var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
        ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
        # Women|Menstr
        bootobs$women_mens[[i]] <- boot(fer_crp %>% filter(Group == "Women|Menstr" & Cohort == "FR97"), statistic = get_ratio_boot, R = boot_n, 
                                   var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
        ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
        # Women|Non-menstr
        bootobs$women_nonmens[[i]] <- boot(fer_crp %>% filter(Group == "Women|Non-menstr" & Cohort == "FR97"), statistic = get_ratio_boot, R = boot_n, 
                                    var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i],var2_trld = CRP_trld, var2_over = T)
        ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
        
        ## Store
        # Men
        means_men[i] <- bootobs$men[[i]]$t0
        upper_men[i] <- ci_obj_men$normal[3]
        lower_men[i] <- ci_obj_men$normal[2]
        # Women|Menstr
        means_women_mens[i] <- bootobs$women_mens[[i]]$t0
        upper_women_mens[i] <- ci_obj_women_mens$normal[3]
        lower_women_mens[i] <- ci_obj_women_mens$normal[2]
        # Women|Non-menstr
        means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
        upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
        lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
        
        # Combine
        means_finrisk <- data.frame(Ferritin = rep(ferritin_values, 3),
                                     means = c(means_men, means_women_mens, means_women_nonmens),
                                     upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                     lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                     Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))
    
    }
    
    # Save
    saveRDS(bootobs, paste0("~/CRP_enrichment/data/PUBL_finrisk_CRP_", boot_n, ".rds"))
} else {
    bootobs <- readRDS(paste0("~/CRP_enrichment/data/PUBL_finrisk_CRP_", boot_n, ".rds"))
    
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    for (i in 1:iterations) {
        ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
        ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
        ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
        # Store
        means_men[i] <- bootobs$men[[i]]$t0
        upper_men[i] <- ci_obj_men$normal[3]
        lower_men[i] <- ci_obj_men$normal[2]
        means_women_mens[i] <- bootobs$women_mens[[i]]$t0
        upper_women_mens[i] <- ci_obj_women_mens$normal[3]
        lower_women_mens[i] <- ci_obj_women_mens$normal[2]
        means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
        upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
        lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
    }
    
    
    # Combine
    means_finrisk <- data.frame(Ferritin = rep(ferritin_values, 3),
                                 means = c(means_men, means_women_mens, means_women_nonmens),
                                 upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                 lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                 Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))
    }

if (!file.exists(paste0("~/CRP_enrichment/data/PUBL_health2k_CRP_", boot_n, ".rds"))) { # run bootstrap only if needed
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    bootobs <- setNames(vector("list", length = 3), c("men", "women_mens", "women_nonmens"))

    for (i in 1:iterations) {
    
    #############
    #### Health2000
    #############
    
    ## Compute
    # Men
    bootobs$men[[i]] <- boot(fer_crp %>% filter(Group == "Men" & Cohort == "H2000"), statistic = get_ratio_boot, R = boot_n, 
                         var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
    ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
    # Women|Menstr
    bootobs$women_mens[[i]] <- boot(fer_crp %>% filter(Group == "Women|Menstr" & Cohort == "H2000"), statistic = get_ratio_boot, R = boot_n, 
                               var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
    ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
    # Women|Non-menstr
    bootobs$women_nonmens[[i]] <- boot(fer_crp %>% filter(Group == "Women|Non-menstr" & Cohort == "H2000"), statistic = get_ratio_boot, R = boot_n, 
                                var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
    ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
    
    ## Store
    # Men
    means_men[i] <- bootobs$men[[i]]$t0
    upper_men[i] <- ci_obj_men$normal[3]
    lower_men[i] <- ci_obj_men$normal[2]
    # Women|Menstr
    means_women_mens[i] <- bootobs$women_mens[[i]]$t0
    upper_women_mens[i] <- ci_obj_women_mens$normal[3]
    lower_women_mens[i] <- ci_obj_women_mens$normal[2]
    # Women|Non-menstr
    means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
    upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
    lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
    
    # Combine
    means_health2k <- data.frame(Ferritin = rep(ferritin_values, 3),
                                 means = c(means_men, means_women_mens, means_women_nonmens),
                                 upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                 lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                 Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))    
    
    
    }
    
    # Save
    saveRDS(bootobs, paste0("~/CRP_enrichment/data/PUBL_health2k_CRP_", boot_n, ".rds"))
} else {
    bootobs <- readRDS(paste0("~/CRP_enrichment/data/PUBL_health2k_CRP_", boot_n, ".rds"))
    
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    for (i in 1:iterations) {
        ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
        ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
        ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
        # Store
        means_men[i] <- bootobs$men[[i]]$t0
        upper_men[i] <- ci_obj_men$normal[3]
        lower_men[i] <- ci_obj_men$normal[2]
        means_women_mens[i] <- bootobs$women_mens[[i]]$t0
        upper_women_mens[i] <- ci_obj_women_mens$normal[3]
        lower_women_mens[i] <- ci_obj_women_mens$normal[2]
        means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
        upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
        lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
    }
    # Combine
    means_health2k <- data.frame(Ferritin = rep(ferritin_values, 3),
                                 means = c(means_men, means_women_mens, means_women_nonmens),
                                 upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                 lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                 Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))
    }

means_all <- rbind(means_finrisk, means_health2k)
means_all$Cohort <- c(rep("FR97", 153), rep("H2000", 153))
means_all$Group <- factor(means_all$Gender, levels = c("Women|Menstr", "Women|Non-menstr", "Men"))
```

## Original analysis plot

```{r ratio_plot_separated}
p <- ggplot(data = means_all, aes(x = Ferritin, y = means)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .3) +
    geom_line(aes(linetype = Group)) +
    theme_minimal() +
    facet_grid(rows = vars(Group), cols = vars(Cohort)) +
    labs(x = expression(paste("Ferritin (", mu, "g/l)")), 
         y = "Change (pp)") + 
    guides(linetype = "none") + 
    theme(text = element_text(size = 15))

# Save
svg("~/CRP_enrichment/results/ratio_plot_separated_rescaled.svg",
    width = 7,
    height = 4.9,
    bg = "transparent")
p
dev.off() -> . # so we avoid output

p
```

Now do the same but strip everyone using OC.

```{r ratio_boot_CRP_by_subgroup_OC}
if (!file.exists(paste0("~/CRP_enrichment/data/PUBL_finrisk_CRP_OC_OMITTED", boot_n, ".rds"))) { # run bootstrap only if needed
    
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    bootobs <- setNames(vector("list", length = 3), c("men", "women_mens", "women_nonmens"))
    
    for (i in 1:iterations) {
        
        #############
        #### FR97
        #############
        
        ## Compute
        # Men
        bootobs$men[[i]] <- boot(fer_crp_omit_OC %>% filter(Group == "Men" & Cohort == "FR97"), statistic = get_ratio_boot, R = boot_n, 
                             var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
        ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
        # Women|Menstr
        bootobs$women_mens[[i]] <- boot(fer_crp_omit_OC %>% filter(Group == "Women|Menstr" & Cohort == "FR97"), statistic = get_ratio_boot, R = boot_n, 
                                   var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
        ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
        # Women|Non-menstr
        bootobs$women_nonmens[[i]] <- boot(fer_crp_omit_OC %>% filter(Group == "Women|Non-menstr" & Cohort == "FR97"), statistic = get_ratio_boot, R = boot_n, 
                                    var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i],var2_trld = CRP_trld, var2_over = T)
        ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
        
        ## Store
        # Men
        means_men[i] <- bootobs$men[[i]]$t0
        upper_men[i] <- ci_obj_men$normal[3]
        lower_men[i] <- ci_obj_men$normal[2]
        # Women|Menstr
        means_women_mens[i] <- bootobs$women_mens[[i]]$t0
        upper_women_mens[i] <- ci_obj_women_mens$normal[3]
        lower_women_mens[i] <- ci_obj_women_mens$normal[2]
        # Women|Non-menstr
        means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
        upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
        lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
        
        # Combine
        means_finrisk <- data.frame(Ferritin = rep(ferritin_values, 3),
                                     means = c(means_men, means_women_mens, means_women_nonmens),
                                     upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                     lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                     Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))
    
    }
    
    # Save
    saveRDS(bootobs, paste0("~/CRP_enrichment/data/PUBL_finrisk_CRP_OC_OMITTED", boot_n, ".rds"))
} else {
    bootobs <- readRDS(paste0("~/CRP_enrichment/data/PUBL_finrisk_CRP_OC_OMITTED", boot_n, ".rds"))
    
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    for (i in 1:iterations) {
        ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
        ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
        ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
        # Store
        means_men[i] <- bootobs$men[[i]]$t0
        upper_men[i] <- ci_obj_men$normal[3]
        lower_men[i] <- ci_obj_men$normal[2]
        means_women_mens[i] <- bootobs$women_mens[[i]]$t0
        upper_women_mens[i] <- ci_obj_women_mens$normal[3]
        lower_women_mens[i] <- ci_obj_women_mens$normal[2]
        means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
        upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
        lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
    }
    
    
    # Combine
    means_finrisk <- data.frame(Ferritin = rep(ferritin_values, 3),
                                 means = c(means_men, means_women_mens, means_women_nonmens),
                                 upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                 lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                 Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))
    }

if (!file.exists(paste0("~/CRP_enrichment/data/PUBL_health2k_CRP_OC_OMITTED", boot_n, ".rds"))) { # run bootstrap only if needed
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    bootobs <- setNames(vector("list", length = 3), c("men", "women_mens", "women_nonmens"))

    for (i in 1:iterations) {
    
    #############
    #### Health2000
    #############
    
    ## Compute
    # Men
    bootobs$men[[i]] <- boot(fer_crp_omit_OC %>% filter(Group == "Men" & Cohort == "H2000"), statistic = get_ratio_boot, R = boot_n, 
                         var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
    ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
    # Women|Menstr
    bootobs$women_mens[[i]] <- boot(fer_crp_omit_OC %>% filter(Group == "Women|Menstr" & Cohort == "H2000"), statistic = get_ratio_boot, R = boot_n, 
                               var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
    ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
    # Women|Non-menstr
    bootobs$women_nonmens[[i]] <- boot(fer_crp_omit_OC %>% filter(Group == "Women|Non-menstr" & Cohort == "H2000"), statistic = get_ratio_boot, R = boot_n, 
                                var1 = Ferritin, var2 = CRP, var1_trld = ferritin_values[i], var2_trld = CRP_trld, var2_over = T)
    ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
    
    ## Store
    # Men
    means_men[i] <- bootobs$men[[i]]$t0
    upper_men[i] <- ci_obj_men$normal[3]
    lower_men[i] <- ci_obj_men$normal[2]
    # Women|Menstr
    means_women_mens[i] <- bootobs$women_mens[[i]]$t0
    upper_women_mens[i] <- ci_obj_women_mens$normal[3]
    lower_women_mens[i] <- ci_obj_women_mens$normal[2]
    # Women|Non-menstr
    means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
    upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
    lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
    
    # Combine
    means_health2k <- data.frame(Ferritin = rep(ferritin_values, 3),
                                 means = c(means_men, means_women_mens, means_women_nonmens),
                                 upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                 lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                 Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))    
    
    
    }
    
    # Save
    saveRDS(bootobs, paste0("~/CRP_enrichment/data/PUBL_health2k_CRP_OC_OMITTED", boot_n, ".rds"))
} else {
    bootobs <- readRDS(paste0("~/CRP_enrichment/data/PUBL_health2k_CRP_OC_OMITTED", boot_n, ".rds"))
    
    ## Preallocate
    # Men
    means_men <- 1:iterations
    upper_men <- 1:iterations
    lower_men <- 1:iterations
    # Women|Menstr
    means_women_mens <- 1:iterations
    upper_women_mens <- 1:iterations
    lower_women_mens <- 1:iterations
    # Women|Non-menstr
    means_women_nonmens <- 1:iterations
    upper_women_nonmens <- 1:iterations
    lower_women_nonmens <- 1:iterations
    
    for (i in 1:iterations) {
        ci_obj_men <- boot.ci(bootobs$men[[i]], type = "norm")
        ci_obj_women_mens <- boot.ci(bootobs$women_mens[[i]], type = "norm")
        ci_obj_women_nonmens <- boot.ci(bootobs$women_nonmens[[i]], type = "norm")
        # Store
        means_men[i] <- bootobs$men[[i]]$t0
        upper_men[i] <- ci_obj_men$normal[3]
        lower_men[i] <- ci_obj_men$normal[2]
        means_women_mens[i] <- bootobs$women_mens[[i]]$t0
        upper_women_mens[i] <- ci_obj_women_mens$normal[3]
        lower_women_mens[i] <- ci_obj_women_mens$normal[2]
        means_women_nonmens[i] <- bootobs$women_nonmens[[i]]$t0
        upper_women_nonmens[i] <- ci_obj_women_nonmens$normal[3]
        lower_women_nonmens[i] <- ci_obj_women_nonmens$normal[2]
    }
    # Combine
    means_health2k <- data.frame(Ferritin = rep(ferritin_values, 3),
                                 means = c(means_men, means_women_mens, means_women_nonmens),
                                 upper = c(upper_men, upper_women_mens, upper_women_nonmens),
                                 lower = c(lower_men, lower_women_mens, lower_women_nonmens),
                                 Gender = c(rep("Men", iterations), rep("Women|Menstr", iterations), rep("Women|Non-menstr", iterations)))
    }

means_all_OC <- rbind(means_finrisk, means_health2k)
means_all_OC$Cohort <- c(rep("FR97", 153), rep("H2000", 153))
means_all_OC$Group <- factor(means_all_OC$Gender, levels = c("Women|Menstr", "Women|Non-menstr", "Men"))
```

# Proportion analysis plot

```{r ratio_plot_separated_OC}
p <- ggplot(data = means_all_OC, aes(x = Ferritin, y = means)) +
    geom_ribbon(aes(ymin = lower, ymax = upper), alpha = .3) +
    geom_line(aes(linetype = Group)) +
    theme_minimal() +
    facet_grid(rows = vars(Group), cols = vars(Cohort)) +
    labs(x = expression(paste("Ferritin (", mu, "g/l)")), 
         y = "Change (pp)") + 
    guides(linetype = "none") + 
    theme(text = element_text(size = 12))

# Save
svg("~/CRP_enrichment/results/ratio_plot_OC_omitted.svg",
    width = 7,
    height = 4.9,
    bg = "transparent")
p
dev.off() -> . # so we avoid output

p
```

What's the diff?

```{r}
comparison_df <- data.frame(Ferritin = means_all$Ferritin,
                            mean = means_all_OC$mean - means_all$mean,
                            lower = means_all_OC$lower - means_all$lower,
                            upper = means_all_OC$upper - means_all$upper,
                            Gender = means_all$Gender,
                            Group = means_all$Group,
                            Cohort = means_all$Cohort)

ggplot(comparison_df, aes(x = Ferritin, y = mean)) + 
    geom_line() +
    facet_grid(rows = vars(Group), cols = vars(Cohort)) +
    theme_minimal()
```

<comment>

How does the fer x crp behave between OC users at the upper end of ferritin?

```{r}
u40 <- fer_crp %>% filter(Group == "Women|Menstr" & Ferritin < 30 & Cohort == "H2000")
o40 <- fer_crp %>% filter(Group == "Women|Menstr" & Ferritin >= 30 & Cohort == "H2000")
u40_OC <- fer_crp_omit_OC %>% filter(Group == "Women|Menstr" & Ferritin < 30 & Cohort == "H2000")
o40_OC <- fer_crp_omit_OC %>% filter(Group == "Women|Menstr" & Ferritin >= 30 & Cohort == "H2000")
CRP_OC_comparison <- bind_rows(u40, o40, u40_OC, o40_OC) %>% mutate(label = c(rep("u30", nrow(u40)),
                                                                              rep("o30", nrow(o40)),
                                                                              rep("u30_OC", nrow(u40_OC)),
                                                                              rep("o30_OC", nrow(o40_OC))))

ggplot(CRP_OC_comparison, aes(x = CRP, group = label, fill = label)) +
    geom_density(alpha = 0.3) +
    #scale_x_log10() +
    theme_minimal()
ggplot(CRP_OC_comparison, aes(x = CRP, group = label, fill = label)) +
    geom_density(alpha = 0.3) +
    scale_x_log10() +
    xlim(0, 4) +
    theme_minimal()
```

# Ratios for OC users

```{r}
h2k_only_oc <- donor_eligible_h2k %>% filter(Hormonal == "Yes")

oc_means <- c()
for (i in 1:iterations){
    oc_means[i] <- get_ratio(h2k_only_oc, Ferritin, CRP, ferritin_values[i], CRP_trld)
}

oc_ratio <- data.frame(Ferritin = ferritin_values, mean = oc_means)

ggplot(oc_ratio, aes(x = Ferritin, y = oc_means)) +
    geom_line() +
    theme_minimal()
```

## What are the ferritin distributions between OC users and non-oc users?

```{r}
non_oc_fer <- de_h2k_omit_OC %>% filter(Group == "Women|Menstr") %>% select(Ferritin)
oc_fer <- h2k_only_oc %>% select(Ferritin)

fer_comparison_df <- data.frame(Ferritin = c(non_oc_fer$Ferritin, oc_fer$Ferritin), Group = c(rep("non", nrow(non_oc_fer)), rep("OC", nrow(oc_fer))))

ggplot(fer_comparison_df, aes(x = Ferritin, group = Group, fill = Group)) +
    geom_density(alpha = 0.3) +
    theme_minimal()
```

# Q: Does the use of OC increase Ferritin more than it does CRP?

## All

```{r}
nonOC <- de_h2k_omit_OC %>% filter(Group == "Women|Menstr") %>% select(Ferritin, CRP)
OC <- h2k_only_oc %>% filter(Group == "Women|Menstr") %>% select(Ferritin, CRP)

summary(nonOC$Ferritin)
summary(OC$Ferritin)
```

```{r}
sum(nonOC$CRP >= 3, na.rm = T) / nrow(nonOC)
sum(OC$CRP >= 3) / nrow(OC)
```

## Fer < 30

```{r}
nonOC <- de_h2k_omit_OC %>% filter(Group == "Women|Menstr" & Ferritin < 30) %>% select(Ferritin, CRP)
OC <- h2k_only_oc %>% filter(Group == "Women|Menstr" & Ferritin < 30) %>% select(Ferritin, CRP)

summary(nonOC$Ferritin)
summary(OC$Ferritin)
```

```{r}
sum(nonOC$CRP >= 3, na.rm = T) / nrow(nonOC)
sum(OC$CRP >= 3) / nrow(OC)
```

## Fer >= 30

```{r}
nonOC <- de_h2k_omit_OC %>% filter(Group == "Women|Menstr" & Ferritin >= 30) %>% select(Ferritin, CRP)
OC <- h2k_only_oc %>% filter(Group == "Women|Menstr" & Ferritin >= 30) %>% select(Ferritin, CRP)

summary(nonOC$Ferritin)
summary(OC$Ferritin)
```

```{r}
sum(nonOC$CRP >= 3, na.rm = T) / nrow(nonOC)
sum(OC$CRP >= 3) / nrow(OC)
```
